Figures

### 2016-02-25 doc/draft.Rmd
updated tables and lofreq

```{r,echo=F}
#Is it clear that I'm only talking about the 10^5 samples until we get to the lower input section
```

@@

WSN33 and PR8 samples were plaque purified and sanger sequenced. cDNA genomes from both viruses were mixed serially such that WSN33 was present at frequencies of 5, 2.5, 1.25, 0.63 and 0.16\% (`r Miseq` A). These viruses differ at 491 positions (primer sites were excluded from analysis), providing 491 true positives in each dilution. On plasmids subjected to limited PCR, DeepSNV identified known variants at 0.1\% frequency with a sensitivity of 0.860 and  specificity of 1.0. Under our experimental conditions and with a slightly more stringent p value  ( p=0.01 vs. p=0.05) we found a moderate reduction in sensitivity  (0.830 for variants at 0.63\% and 0.14 for variants at 0.16\%). The specificity was above 0.998 in all dilutions. While this drop in specificity appears minor, it is important to keep in mind that even a false positive rate of 0.002 results in 78 false positives when applied across the `r miseq_vars` potential variants present in the `r miseq_kb`kb of the influenza genome we analyzed (3 variants per genomic position).

@@

The samples with the greatest input concentration and variant frequencies of 1\% and greater sensitivities of 0.850 or greater (`r Hiseq` B and C). Unlike what was seen in cell culture samples abover, sensitivity dropped drastically at 0.5\% frequency. Typically our samples grown in cell culture have  at least $10^8$ genomes per microliter; whereas, the nucleic acid concentration in these samples was lower by 1,000 fold.  The reduced sensitivity is likely due to the fact that we expect only 7,000 copies of each variant genome in the 0.5\% frequency samples compared to the equivalent of 62,500,000 variant genomes in the 0.63\% sample from the cell culture experiment above. During library preperation, a disproportionate amount of the 2.0\% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0\% sample.

@@

In addition to distinguishing true from false SNV, DeepSNV was able to estimate the frequency of the true positive variants (`r Freq`). Although the median of the measured frequencies match the expetect values we found substantial spread about the median and the overall the fit was quite poor ( R^2^=`r corr_coef`). On average the percent difference between the measured frequency an the expected was `r mean_fold`\%. This error should be kept in mind when employing down stream analysis that depend on frequency measurements such as diversity metrics like Shannon's Entropy and Simpson's D, as well as inter-sample comparison of population structure .

@@

DeepSNV is one of many variant callers that use a combination of empiric and statistical approaches to model error rates and separate true from false variants. We asked whether the decreased accuracy observed in our data set was due simply to peculiarities specific to the DeepSNV algorithm. We therefore analyzed our $10^5$ input populations using Lofreq, another algorithm that is commonly used in viral next generation sequencing studies. Lofreq has been reported to have perfect specificity in the past. Under our conditions Lofreq had marginally reduced sensitivity compared to DeepSNV when applied to the variant frequencies $\geq$ 1.0\% but marginally increased sensitivity when applied to variant frequencies $<$ 1.0\% (`r Lofreq`). The specificity of LoFreq was comparable to what we observed with DeepSNV in our high-input cell culture-derived populations (`r Miseq`), and better than DeepSNV in our initial implementation (`r Hiseq` prior to Phred, MapQ, and read position filtering). This increased specificity is most likely due to the fast that  the LoFreq algorithm takes MapQ and Phred scores into account when calling variants and has a stringent strand bias filter that removes many of the variants fount only at one end of the reads. However even with these additional characteristics, the specificity of LoFreq was lower than our improved DeepSNV pipeline (`r Qual`) with over 40 false positives per sample. It appears that higher than expected false positive rates are not specific to DeepSNV and most likely plague all variant callers applied to patient-derived viral samples.

%%%% if the Errors are PCR related why does Lofreq remove them.

@@

Host-derived viral populations vary in copy number and titer by several orders of magnitude. This variability can be attributed to a variety of factors including but not limited to: collection site, ease of of nucleic acid isolation, library preparation pipeline, and host and viral variability. To ensure accuracy across a range of input levels, we diluted our experimental populations serially into basal media and and called variants using our modified DeepSNV analysis pipeline (`r Lower`). As expected, the sensitivity was lower in populations with fewer genomes. A variant with a 0.5\% frequency in a $10^4$ genomes per microliter sample is expected to present on only 700 genomes in the initial RT-PCR and any number of these genomes could be lost due to bottlenecks in the amplification and library preparation process. We also found that specificity was reduced in samples with lower starting input. This is presumably due to a greater dependence on RT-PCR amplification. These data highlight the importance of controlling for input levels when comparing diversity across experimental samples.

In most cases, RT-PCRs error should be randomly distributed across the amplified region. If RT-PCR errors are responsible for the reduced specificity found at lower input levels, and RT-PCR are randomly distributed across the genome, we would expect such error to be eliminated by processing the samples in duplicate. To test this hypothesis, we performed duplicate RT-PCR reactions on the 5, 2, 1, and 0.5\% variant frequency samples that had a starting concentration of $10^4$ genomes per microliter. This was done using a separate RNA prep than that shown in `r lower` A. The duplicates were processed separately, but sequenced on the same lane of an Illumina Hiseq. We applied the previously mentioned quality cut-offs and required that a given variant be found in both duplicates. This drastically improved our specificity and resulted in an accuracy that was comparable to that found in the $10^5$ samples (`r Dups`).

>>>>

```{r,echo=F}
#Is it clear that I'm only talking about the 10^5 samples until we get to the lower input section
# verfiy # of Tp in 0.5% 10^5 sample
```

@@

WSN33 and PR8 samples were plaque purified and Sanger sequenced. cDNA genomes from both viruses were mixed serially such that WSN33 was present at frequencies of 5, 2.5, 1.25, 0.63 and 0.16\% (`r Miseq` A). These viruses differ at 491 positions (primer sites were excluded from analysis), providing 491 true positives in each dilution. On plasmids subjected to limited PCR, DeepSNV identified known variants at 0.1\% frequency with a sensitivity of 0.860 and  specificity of 1.0. Under our experimental conditions and with a slightly more stringent p value  ( p=0.01 vs. p=0.05) we found a moderate reduction in sensitivity  (0.830 for variants at 0.63\% and 0.14 for variants at 0.16\%). The specificity was above 0.998 in all dilutions. While this drop in specificity appears minor, it is important to keep in mind that even a false positive rate of 0.002 results in 78 false positives when applied across the `r miseq_vars` potential variants present in the `r miseq_kb`kb of the influenza genome we analyzed (3 variants per genomic position).

@@

The samples with the greatest input concentration and variant frequencies of 1\% and greater sensitivities of 0.850 or greater (`r Hiseq` B and C). Unlike what was seen in cell culture samples above, sensitivity dropped drastically at 0.5\% frequency. Typically our samples grown in cell culture have  at least $10^8$ genomes per microliter; whereas, the nucleic acid concentration in these samples was lower by 1,000 fold.  The reduced sensitivity is likely due to the fact that we expect only 7,000 copies of each variant genome in the 0.5\% frequency samples compared to the equivalent of 62,500,000 variant genomes in the 0.63\% sample from the cell culture experiment above. During library preparation, a disproportionate amount of the 2.0\% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0\% sample.

@@

In addition to distinguishing true from false SNV, DeepSNV was able to estimate the frequency of the true positive variants (`r Freq`). Although the median of the measured frequencies match the expect values, we found substantial spread in each sample and the overall the fit was quite poor ( R^2^=`r corr_coef`). The mean percent difference between the measured frequency an the expected was `r mean_fold`\%. This error should be kept in mind when employing down stream analysis that depend on frequency measurements such as diversity metrics like Shannon's Entropy and Simpson's D, as well as inter-sample comparison of population structure .

@@

DeepSNV is one of many variant callers that use a combination of empiric and statistical approaches to model error rates and separate true from false variants. We asked whether the decreased accuracy observed in our data set was due simply to peculiarities specific to  DeepSNV. We therefore analyzed our $10^5$ input populations using Lofreq, another algorithm that is commonly used in viral next generation sequencing studies. Lofreq has been reported to have perfect specificity in the past. Under our conditions Lofreq had marginally reduced sensitivity compared to DeepSNV when applied to the variant frequencies $\geq$ 1.0\% but marginally increased sensitivity when applied to variant frequencies $<$ 1.0\% (`r Lofreq`). The specificity of LoFreq was comparable to what we observed with DeepSNV in our high-input cell culture-derived populations (`r Miseq`), and better than DeepSNV in our initial implementation (`r Hiseq` prior to Phred, MapQ, and read position filtering). This increased specificity is most likely due to the fast that  the LoFreq algorithm takes MapQ and Phred scores into account when calling variants and has a stringent strand bias filter that removes many of the variants fount only at one end of the reads. However even with these additional characteristics, the specificity of LoFreq was lower than our improved DeepSNV pipeline (`r Qual`) with over 40 false positives per sample. It appears that higher than expected false positive rates are not specific to DeepSNV and most likely plague all variant callers applied to patient-derived viral samples.

@@

Host-derived viral populations vary in copy number and titer by several orders of magnitude. This variability can be attributed to a variety of factors including but not limited to: collection site, ease of of nucleic acid isolation, library preparation, and host and viral variability. To ensure accuracy across a range of input levels, we diluted our experimental populations serially into basal media and and identified variants using our modified DeepSNV analysis pipeline (`r Lower`). As expected, the sensitivity was lower in populations with fewer genomes. A variant with a 0.5\% frequency in a $10^4$ genomes per microliter sample is expected to present on only 700 genomes in the initial RT-PCR and it is expected that many of these genomes will be lost due to bottlenecks in the amplification and library preparation process. We also found that specificity was reduced in samples with lower starting input. This is presumably due to a greater dependence on RT-PCR amplification. These data highlight the importance of controlling for input levels when comparing diversity across experimental samples.

In most cases, RT-PCRs error should be randomly distributed across the amplified region. If RT-PCR errors are responsible for the reduced specificity found at lower input levels, and RT-PCR are randomly distributed across the genome, we would expect such error to be eliminated by processing the samples in duplicate. To test this hypothesis, we performed duplicate RT-PCR reactions on the 5, 2, 1, and 0.5\% variant frequency samples that had a starting concentration of $10^4$ genomes per microliter. This was done using a separate RNA prep than that shown in `r lower` A. The duplicates were processed separately, but sequenced on the same lane of an Illumina Hiseq. We applied the previously mentioned quality cut-offs and required that a given variant be found in both duplicates. This drastically improved our specificity and resulted in an accuracy that was comparable to that found in the $10^5$ samples (`r Dups`). Thus accurate analysis of low input RNA virus samples can be achieved through duplicate RT-PCR and careful benchmarking experiments.

####The consequences of inaccurate SNV identification

`r Table` stresses the importance of accurate SNV indentification in studies of viral diversity. We have apply three commonly used diversity metrics, (Richness, Shannon's Entropy, the inverse of Simpson's D), to our $10^5$ genomes per microliter populations.

### 2016-02-25 scripts/figures.Rmd
updated tables and lofreq

pdf("../results/figures/roc.example.pdf",width=3.4,height=4)
plot.ex(0.35,0.95,roc.df,example)
dev.off()
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/roc.ex.pdf",width = 12,height = 8)
# plot.ex(0.35,0.95,roc.df,example)
# dev.off()

@@

  if("gc" %in% names(x)){
    y$gc=x$gc[1]
  }
  return(y)
}

sum_roc<-function(x){
  sum.df<-subset(x,category %in% c(TRUE,FALSE),select=c(category,p.val))# get the TRUE and false variant calls and p.vals
  if(length(which(sum.df$category==T))>0){ # filter out cases where there aren't any TP found
    if(length(which(sum.df$category==F))==0){
      sum.df<-rbind(sum.df,data.frame(category=F,p.val=1))
    }  
    roc(sum.df$category~sum.df$p.val,plot=F,CI=T,direction=">")
  }
}

@@

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
} 
######### Miseq roc functions ##########
miseq.roc<-function(sum.df,possible_tp,possible_vars){ # A function to run the roc calculations and adjustments
  roc.ls<-dlply(sum.df,~Id,sum_roc)

@@

  roc.df.adj<-mutate(roc.df.adj,exp.freq=as.factor(find_freq.v(samp)))
  #fills<-ddply(roc.df.adj,~samp,fill_in_plots) # take out these lines to not fill in the plots
  #roc.df.adj<-rbind(fills,roc.df.adj) # here too
  roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
  roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
  return(roc.df.adj)
}
miseq.roc.table<-function(sum.df,cut.off,possible_tp,possible_vars){
  roc.ls<-dlply(sum.df,~Id,sum_roc)
  roc.df<-roc_df.one(roc.ls,cut.off)
  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
  roc.df.adj<-roc.df.adj[order(roc.df.adj$exp.freq,decreasing = T),]
  roc.table<-subset(roc.df.adj,select=c(exp.freq,adj.sensitivity,TP,adj.specificity,FP))
  return(roc.table)
}

@@

######### Hiseq roc functions ##########
hiseq.roc<-function(sum.df,possible_tp,possible_vars){ # A function to run the roc calculations and adjustments
  roc.ls<-dlply(sum.df,~Id,sum_roc)

@@

hiseq.roc.table<-function(sum.df,cut.off,possible_tp,possible_vars){
  roc.ls<-dlply(sum.df,~Id,sum_roc)
  roc.df<-roc_df.one(roc.ls,cut.off)
  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
mutate(roc.df.adj, gc = as.numeric(sub(pat,"\\1",samp)),
       exp.freq = sub(pat,"\\2",samp)
)-> roc.df.adj

@@

@@

m.roc.df<-miseq.roc(sum.df,possible_tp,possible_vars)

@@

m.roc.table<-miseq.roc.table(sum.df,0.01,possible_tp,possible_vars)

@@

pdf("../results/figures/miseq1.pdf",width=3.4,height=7)
ggdraw() +
  draw_plot(miseq.setup, 0.2, .66, .6, 0.33) + #x y width height
  draw_plot(miseq.roc.p, 0.05, 0.33, .9, 0.33) +
  draw_plot(tbl, 0.05, 0, .9, 0.33) +
  draw_plot_label(c("A", "B", "C"), c(0.01, 0.01, 0.01), c(0.95, 0.7, 0.35), size = 15)
dev.off()

@@

```{r,Miseq_read.cut,eval=F}
### With read position cut off 63 and 188
read.df<-subset(sum.df,Read_pos>63 & Read_pos<188)
m.roc.df<-miseq.roc(read.df,possible_tp,possible_vars)

@@

m.roc.table<-miseq.roc.table(read.df,0.01,possible_tp,possible_vars)
knitr::kable(m.roc.table)

@@

h.roc.df<-hiseq.roc(sum.df,possible_tp,possible_vars)

@@

h.table.df<-hiseq.roc.table(sum.df,0.01,possible_tp,possible_vars)

@@

pdf("../results/figures/hiseq1.pdf",width=3.4,height=7)
ggdraw() +
  draw_plot(hiseq.setup, 0.1, .57, 0.8, 0.43) + #x y width height
  draw_plot(Hiseq.roc, 0.05, 0.33, .8, 0.33) +
  draw_plot(tbl, 0.05, 0, .9, 0.33) +
  draw_plot_label(c("A", "B", "C"), c(0.01, 0.01, 0.01), c(1, 0.7, 0.35), size = 15)
dev.off()

@@

cut.df<-subset(sum.df,MapQ>30 & Phred>35) # This removes the variants that don't pass

@@

sig.cut.df<-subset(cut.df,p.val<0.01)

@@

@@

h.cut.read.df<-hiseq.roc(cut.read.df.all,possible_tp,possible_vars)

@@

h.cut.read.table<-hiseq.roc.table(cut.read.df.all,0.01,possible_tp,possible_vars)
#knitr::kable(subset(h.cut.read.table,gc==5))

@@

pdf("../results/figures/figure4.pdf",width=3.4,height=9)
ggdraw() +
  draw_plot(mapq_phred.normal,0.03, 0.77, .85, 0.23) + 
  draw_plot(read.pos, 0.03, .53, .85, .25) +
  draw_plot(read.roc, 0.03, 0.23, 0.85, .3) +
  draw_plot(tbl,   0.025, 0, 0.95, .25) +
  draw_plot_label(c("A", "B", "C","D"), c(0, 0, 0,0), c(1, 0.78, 0.55,0.25), size = 15)
dev.off()

@@

lofreq.df<-read.csv("../data/process/2015-6-23/lofreq/all.removed.mapq_vcf.vcf_csv.csv")

@@

lofreq.roc<-hiseq.roc(cut.df,possible_tp,possible_vars)

@@

@@

x$adj.sensitivity[x$frequency=="1.0%"]<-x$adj.sensitivity[x$frequency=="1.0%"]-0.04
x$adj.sensitivity[x$frequency=="2.0%"]<-x$adj.sensitivity[x$frequency=="2.0%"]-0.12

@@

pdf("../results/figures/lofreq1.pdf",width=3.4,height=5)
ggdraw() +
  draw_plot(lofreq.roc.p, 0, 0.5, 0.9, 0.45) + 
  draw_plot(tbl, 0.025, 0, 0.95, .48) +
  draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)
dev.off()

ggdraw() +
  draw_plot(lofreq.roc.p, 0, 0.5, 0.9, 0.45) + 
  draw_plot(tbl, 0.025, 0, 0.95, .48) +
  draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)

@@

@@

pdf("../results/figures/freq.pdf",width=3.4,height=3.4)
ac_freq
dev.off()

@@

pdf("../results/figures/lower.pdf",width=3.4,height=9)
ggdraw() +
  draw_plot(roc.4,0.03, 0.69, .85, 0.25) + 
  draw_plot(tbl.4, 0.025, .47,  0.95, .25) +
  draw_plot(roc.3, 0.03, 0.22, 0.85, .25) +
  draw_plot(tbl.3,   0.025, 0, 0.95, .25) +
  draw_plot_label(c("A", "B", "C","D"), c(0, 0, 0,0), c(0.97, 0.73, 0.5,0.23), size = 15)
dev.off()

@@

dups.roc<-hiseq.roc(dups,possible_tp,possible_vars)

@@

dups.table<-hiseq.roc.table(dups,0.01,possible_tp,possible_vars)
#knitr::kable(subset(dups.table))

@@

pdf("../results/figures/dups.roc.pdf",width=3.4,height=5)
ggdraw() +
  draw_plot(dups.roc.p, 0, 0.5, 0.9, 0.45) + 
  draw_plot(tbl, 0.025, 0, 0.95, .48) +
  draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)
dev.off()

@@

```{r}
perfect<-data.frame(mutation=rep(true_snv$mutant,times=4),freq.var=rep(c(0.05,0.02,0.01,0.005),each=20))
dups<-subset(dups,p.val<0.01)
dups<-mutate(dups,freq.var=pmax(freq1,freq2))

@@

H<-function(x){
  p=x$freq.var*log(x$freq.var)
  -sum(p)
}

D<-function(x){
  p=(x$freq.var)^2
  1/sum(p)
}

euc<-function(x,ref){
  x$Id="test"
  ref$Id="ref"
  all<-rbind(subset(x,select=c(mutation,freq.var,Id)),ref)
  tab<-as.matrix(subset(dcast(all,mutation ~ Id, value.var = "freq.var"),select=-c(mutation)))
  tab[is.na(tab)]<-0
  as.dist(dist(tab))
}
get_data<-function(x,setting){
  if(length(unique(x$freq.var))==1){
    exp.freq=unique(x$freq.var)
  }else{
  exp.freq=unique(x$exp.freq)
  }
  out<-data.frame(Richness=rich(x), Entropy=H(x),Simpson=D(x))#,dist=euc(x,subset(perfect,freq.var==exp.freq))[1])
  rownames(out)<-setting
  return(t(out))
}
tabulate<-function(input,freq){
get_data(subset(perfect,freq.var==freq),"Perfect")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq),"Lofreq")->lo
get_data(subset(cut.read.sig.df,gc==input & exp.freq==freq),"DeepSNV Final")->deepSNV.f
knitr::kable(cbind(perf,deepSNV,lo,deepSNV.f),digits = 2)
}
tabulate.4<-function(input,freq){
get_data(subset(perfect,freq.var==freq),"Perfect")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq),"Lofreq")->lo
get_data(subset(cut.read.sig.df,gc==input & exp.freq==freq),"DeepSNV Final")->deepSNV.f
get_data(subset(dups,exp.freq==freq),"DeepSNV duplicates")->deepSNV.d

knitr::kable(cbind(perf,deepSNV,lo,deepSNV.f,deepSNV.d),digits = 2)
}

@@

$10^4$ 5%

```{r}
tabulate.4(4,0.05)
```

$10^4$ 1%

```{r}
tabulate.4(4,0.01)
```

$10^4$ 0.5%

```{r}
tabulate.4(4,0.005)
```

>>>>

# pdf("../results/figures/roc.example.pdf",width=3.4,height=4)
# plot.ex(0.35,0.95,roc.df,example)
# dev.off()
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/roc.ex.pdf",width = 12,height = 8)
# plot.ex(0.35,0.95,roc.df,example)
# dev.off()

@@

  if("gc" %in% names(x)){
    y$gc=x$gc[1]
  }
  return(y)
}
sum_roc<-function(x,direction){
  sum.df<-subset(x,category %in% c(TRUE,FALSE),select=c(category,p.val))# get the TRUE and false variant calls and p.vals
  if(length(which(sum.df$category==T))>0){ # filter out cases where there aren't any TP found
    if(length(which(sum.df$category==F))==0){
      sum.df<-rbind(sum.df,data.frame(category=F,p.val=1))
    }  
    roc(sum.df$category~sum.df$p.val,plot=F,CI=T,direction=direction)
  }
}

@@

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
} 
######### Miseq roc functions ##########
miseq.roc<-function(sum.df,possible_tp,possible_vars,direction){ # A function to run the roc calculations and adjustments
  roc.ls<-dlply(sum.df,~Id,sum_roc,direction)

@@

  roc.df.adj<-mutate(roc.df.adj,exp.freq=as.factor(find_freq.v(samp)))
  #fills<-ddply(roc.df.adj,~samp,fill_in_plots) # take out these lines to not fill in the plots
  #roc.df.adj<-rbind(fills,roc.df.adj) # here too
  roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
  roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
  return(roc.df.adj)
}
miseq.roc.table<-function(sum.df,cut.off,possible_tp,possible_vars,direction){
  roc.ls<-dlply(sum.df,~Id,sum_roc,direction)
  roc.df<-roc_df.one(roc.ls,cut.off)
  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
  roc.df.adj<-roc.df.adj[order(roc.df.adj$exp.freq,decreasing = T),]
  roc.table<-subset(roc.df.adj,select=c(exp.freq,adj.sensitivity,TP,adj.specificity,FP))
  return(roc.table)
}

@@

######### Hiseq roc functions ##########
hiseq.roc<-function(sum.df,possible_tp,possible_vars,direction){ # A function to run the roc calculations and adjustments
  roc.ls<-dlply(sum.df,~Id,sum_roc,direction)

@@

hiseq.roc.table<-function(sum.df,cut.off,possible_tp,possible_vars,direction){
  roc.ls<-dlply(sum.df,~Id,sum_roc,direction)
  roc.df<-roc_df.one(roc.ls,cut.off)
  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
mutate(roc.df.adj, gc = as.numeric(sub(pat,"\\1",samp)),
       exp.freq = sub(pat,"\\2",samp)
)-> roc.df.adj

@@

possible_vars<-sum(regions.bed$length)*3 # three possible variants/ position
miseq_vars<-sum(regions.bed$length)*3
miseq_kb<-sum(regions.bed$length)/1000

@@

m.roc.df<-miseq.roc(sum.df,possible_tp,possible_vars,">")

@@

m.roc.table<-miseq.roc.table(sum.df,0.01,possible_tp,possible_vars,">")

@@

# pdf("../results/figures/miseq1.pdf",width=3.4,height=7)
# ggdraw() +
#   draw_plot(miseq.setup, 0.2, .66, .6, 0.33) + #x y width height
#   draw_plot(miseq.roc.p, 0.05, 0.33, .9, 0.33) +
#   draw_plot(tbl, 0.05, 0, .9, 0.33) +
#   draw_plot_label(c("A", "B", "C"), c(0.01, 0.01, 0.01), c(0.95, 0.7, 0.35), size = 15)
# dev.off()

@@

```{r,Miseq_read.cut,eval=F}
### With read position cut off 63 and 188
read.df<-subset(sum.df,Read_pos>63 & Read_pos<188)
m.roc.df<-miseq.roc(read.df,possible_tp,possible_vars,">")

@@

m.roc.table<-miseq.roc.table(read.df,0.01,possible_tp,possible_vars,">")
knitr::kable(m.roc.table)

@@

hiseq_vars<-possible_vars
hiseq_kb<-sum(regions.bed$length)/1000

h.roc.df<-hiseq.roc(sum.df,possible_tp,possible_vars,">")

@@

h.table.df<-hiseq.roc.table(sum.df,0.01,possible_tp,possible_vars,">")

@@

# pdf("../results/figures/hiseq1.pdf",width=3.4,height=7)
# ggdraw() +
#   draw_plot(hiseq.setup, 0.1, .57, 0.8, 0.43) + #x y width height
#   draw_plot(Hiseq.roc, 0.05, 0.33, .8, 0.33) +
#   draw_plot(tbl, 0.05, 0, .9, 0.33) +
#   draw_plot_label(c("A", "B", "C"), c(0.01, 0.01, 0.01), c(1, 0.7, 0.35), size = 15)
# dev.off()

@@

#cut.df<-subset(sum.df,MapQ>30 & Phred>35) # This removes the variants that don't pass

@@

sig.cut.df<-subset(cut.df.all,p.val<0.01)

@@

fp_ends<-subset(sig.cut.df, category==F & Read_pos<32 | Read_pos>94)
false_p_phred<-mean(fp_ends$Phred)

@@

h.cut.read.df<-hiseq.roc(cut.read.df.all,possible_tp,possible_vars,">")

@@

h.cut.read.table<-hiseq.roc.table(cut.read.df.all,0.01,possible_tp,possible_vars,">")
#knitr::kable(subset(h.cut.read.table,gc==5))

@@

# pdf("../results/figures/figure4.pdf",width=3.4,height=9)
# ggdraw() +
#   draw_plot(mapq_phred.normal,0.03, 0.77, .85, 0.23) + 
#   draw_plot(read.pos, 0.03, .53, .85, .25) +
#   draw_plot(read.roc, 0.03, 0.23, 0.85, .3) +
#   draw_plot(tbl,   0.025, 0, 0.95, .25) +
#   draw_plot_label(c("A", "B", "C","D"), c(0, 0, 0,0), c(1, 0.78, 0.55,0.25), size = 15)
# dev.off()

@@

lofreq.df<-read.csv("../data/process/2015-6-23/lofreq/all.removed.mapq_vcf.vcf_csv.csv",stringsAsFactors = F)

@@

direction<-">"
lofreq.roc<-hiseq.roc(lofreq.df,possible_tp,possible_vars,"<")

@@

x<-mutate(x,frequency=as.character(find_freq.v.h(exp.freq)))

@@

x$adj.sensitivity[x$frequency=="1.0%"]<-x$adj.sensitivity[x$frequency=="1.0%"]+0.02
x$adj.sensitivity[x$frequency=="2.0%"]<-x$adj.sensitivity[x$frequency=="2.0%"]-0.02

@@

# pdf("../results/figures/lofreq1.pdf",width=3.4,height=5)
# ggdraw() +
#   draw_plot(lofreq.roc.p, 0, 0.5, 0.9, 0.45) + 
#   draw_plot(tbl, 0.025, 0, 0.95, .48) +
#   draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)
# dev.off()

ggdraw() +
  draw_plot(lofreq.roc.p, 0, 0.5, 0.9, 0.45) + 
  draw_plot(tbl, 0.025, 0, 0.95, .48) +
  draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)
direction<-"<"
```

@@

### here weget the R2 of the sample

model<-lm(formula = freq.var~exp.freq,data = subset(cut.read.sig.df,category==T & gc=="5"))

corr_coef<-summary(model)$adj.r.squared

interest<-subset(cut.read.sig.df,category==T & gc=="5")

mean_fold<-mean(abs(interest$freq.var-interest$exp.freq)/interest$exp.freq)

print(mean_fold)

@@

# pdf("../results/figures/freq.pdf",width=3.4,height=3.4)
# ac_freq
# dev.off()

@@

# pdf("../results/figures/lower.pdf",width=3.4,height=9)
# ggdraw() +
#   draw_plot(roc.4,0.03, 0.69, .85, 0.25) + 
#   draw_plot(tbl.4, 0.025, .47,  0.95, .25) +
#   draw_plot(roc.3, 0.03, 0.22, 0.85, .25) +
#   draw_plot(tbl.3,   0.025, 0, 0.95, .25) +
#   draw_plot_label(c("A", "B", "C","D"), c(0, 0, 0,0), c(0.97, 0.73, 0.5,0.23), size = 15)
# dev.off()

@@

dups.roc<-hiseq.roc(dups,possible_tp,possible_vars,">")

@@

dups.table<-hiseq.roc.table(dups,0.01,possible_tp,possible_vars,">")
#knitr::kable(subset(dups.table))

@@

# pdf("../results/figures/dups.roc.pdf",width=3.4,height=5)
# ggdraw() +
#   draw_plot(dups.roc.p, 0, 0.5, 0.9, 0.45) + 
#   draw_plot(tbl, 0.025, 0, 0.95, .48) +
#   draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)
# dev.off()

@@

```{r}
mut_pat="(.*)_[A,T,C,G](\\d+)[A,T,C,G]"
perfect<-data.frame(mutation=rep(true_snv$mutant,times=4),freq.var=rep(c(0.05,0.02,0.01,0.005),each=20))
perfect=mutate(perfect,chr=gsub(pattern = mut_pat,replacement = "\\1",x=mutation),pos=gsub(pattern = mut_pat,replacement = "\\2",x=mutation))
dups<-subset(dups,p.val<0.01)
dups<-mutate(dups,freq.var=pmax(freq1,freq2))

@@

H<-function(x){
  H_pos<-ddply(x,~chr+pos,summarize,wt=1-sum(freq.var),H=-sum(c(freq.var*log(freq.var),wt*log(wt))))
sum(H_pos$H)/(possible_vars/3)
}

D<-function(x){
  D_pos<-ddply(x,~chr+pos,summarize,wt=1-sum(freq.var),D=1/sum(c(freq.var^2,wt^2)))
  others=(possible_vars/3)-dim(D_pos)[1]
  sum(D_pos$D,others)/(possible_vars/3)
}

euc<-function(x,ref){
  x$Id="test"
  ref$Id="ref"
  all<-rbind(subset(x,select=c(mutation,freq.var,Id)),subset(ref,select=c(mutation,freq.var,Id)))
  tab<-as.matrix(subset(dcast(all,mutation ~ Id, value.var = "freq.var"),select=-c(mutation)))
  tab[is.na(tab)]<-0
  dist(t(tab),"manhattan")[1]
}
get_data<-function(x,setting){
  if(length(unique(x$freq.var))==1){
    exp.freq=unique(x$freq.var) # this is the perfect condition where there is only 1 frequency and it is all the same
  }else{
  exp.freq=unique(x$exp.freq)
  }
  out<-data.frame(Richness=round(rich(x),digits = 0), Entropy=format(H(x),scientific = T,digits = 3),"1/D"=D(x),Manhattan=round(euc(x,subset(perfect,freq.var==exp.freq)),digits=3))
  rownames(out)<-setting
  return(t(out))
}
tabulate<-function(input,freq){
get_data(subset(perfect,freq.var==freq),"Perfect")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq & p.val<0.01),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq),"Lofreq")->lo
get_data(subset(cut.read.df.all,gc==input & exp.freq==freq & p.val<0.01),"DeepSNV Final")->deepSNV.f
knitr::kable(cbind(perf,deepSNV,lo,deepSNV.f))
}
# tabulate.4<-function(input,freq){
# get_data(subset(perfect,freq.var==freq),"Perfect")->perf
# get_data(subset(sum.df,gc==input & exp.freq==freq),"DeepSNV")->deepSNV
# get_data(subset(lofreq.df,gc==input & exp.freq==freq),"Lofreq")->lo
# get_data(subset(cut.read.sig.df,gc==input & exp.freq==freq),"DeepSNV Final")->deepSNV.f
# get_data(subset(dups,exp.freq==freq),"DeepSNV duplicates")->deepSNV.d
# 
# knitr::kable(cbind(perf,deepSNV,lo,deepSNV.f,deepSNV.d),digits = 2)
# }

@@
